TrenchR: an R package for transparent environmental and ecological biophysics

### 2022-05-06 ManuscriptTe.Rmd
Ms updates

##Calculate operative temperature, Te
dat$Te<-Tb_grasshopper(T_a=dat$Tgrass, T_g=dat$SoilTemp, u=dat$Ugrass, H=dat$Rad, K_t=1, psi=dat$psi, L=0.0211, Acondfact=0.0, abs=0.9, r_g=0.5)

#T_a	and T_g are air and surface temperatures (C), respectively; u	is wind speed (m/s); 
#H is total (direct + diffuse) solar radiation flux (W/m^2); 
#K_t	is the clearness index (dimensionless), which is the ratio of the global solar radiation measured at the surface to the total solar radiation at the top of the atmosphere;
#psi is solar zenith angle (degrees); L	is grasshopper length (m);
#Acondfact is the proportion of the grasshopper surface area in contact with the ground;
#abs is absorptivity of grasshopper to solar radiation (proportion);
#r_g is substrate solar reflectivity (proportion).
```

@@

We then implement a generic energy budget. We first illustrate the estimation of convective heat transfer coefficients as a function of wind speed.
```{r}
#Estimate heat transfer coefficient using empirical relationship
dat$H_L=heat_transfer_coefficient(V=dat$Ugrass,D=0.0211,K=0.025, nu= 15.3 * 10^(-6) , taxa="cylinder")
#where V is	air velocity (m/s); D is the characteristic dimension (m);
#K is the thermal conductivity of air (W m^-1 K^-1);
#nu is the kinematic viscosity of air (m^2 s^-1)

#Also estimate using a spherical approximation and a simplified version of the approximation.
H_L1= heat_transfer_coefficient_approximation(V=dat$Ugrass,D=0.0211,K= 0.025, nu= 15.3 * 10^(-6), taxa="flyinginsect")
H_L2= heat_transfer_coefficient_simple(V=dat$Ugrass,D=0.0211)

#surface area from length(m)
A= sa_from_length(l=0.0211)
```

>>>>

##Calculate operative temperature, Te
dat$Te<-Tb_grasshopper(T_a=dat$Tgrass, T_g=dat$SoilTemp, u=dat$Ugrass, H=dat$Rad, K_t=1, psi=dat$psi, l=0.0211, Acondfact=0.0, abs=0.9, r_g=0.5)

#T_a	and T_g are air and surface temperatures (C), respectively; u	is wind speed (m/s); 
#H is total (direct + diffuse) solar radiation flux (W/m^2); 
#K_t	is the clearness index (dimensionless), which is the ratio of the global solar radiation measured at the surface to the total solar radiation at the top of the atmosphere;
#psi is solar zenith angle (degrees); l	is grasshopper length (m);
#Acondfact is the proportion of the grasshopper surface area in contact with the ground;
#abs is absorptivity of grasshopper to solar radiation (proportion);
#r_g is substrate solar reflectivity (proportion).
```

@@

We then implement a generic energy budget. We first illustrate the estimation of convective heat transfer coefficients as a function of wind speed.
```{r}
#Estimate heat transfer coefficient using empirical relationship
dat$H_L=heat_transfer_coefficient(V=dat$Ugrass,D=0.0211,K=0.025, nu= 15.3 * 10^(-6) , taxon="cylinder")
#where V is	air velocity (m/s); D is the characteristic dimension (m);
#K is the thermal conductivity of air (W m^-1 K^-1);
#nu is the kinematic viscosity of air (m^2 s^-1)

#Also estimate using a spherical approximation and a simplified version of the approximation.
H_L1= heat_transfer_coefficient_approximation(V=dat$Ugrass,D=0.0211,K= 0.025, nu= 15.3 * 10^(-6), taxon="flyinginsect")
H_L2= heat_transfer_coefficient_simple(V=dat$Ugrass,D=0.0211, type="Gates")

#surface area from length(m)
A= surface_area_from_length(l=0.0211)
```

### 2022-05-06 Manuscript_full.Rmd
Ms updates

---
title: 'TrenchR: an R package for transparent environmental and ecological biophysics'
author: "Lauren B. Buckley, Bryan A. Briones Ortiz, Isaac Caruso, Aji John, Ofir Levy, Abigail V. Meyer, Eric A. Riddell, and Yutaro Sakairi"
output:
  html_document: default
  word_document: default
  pdf_document: default
bibliography: bibliography.bib
csl: methods-in-ecology-and-evolution.csl
editor_options: 
  markdown: 
    wrap: 72
---

@@

Lauren B. Buckley^1^, Bryan A. Briones Ortiz^1^, Isaac Caruso^1^, Aji
John^1^, Ofir Levy^2^, Abigail V. Meyer^1^, Eric A. Riddell^3^, Yutaro Sakairi^1^

1.  Department of Biology, University of Washington, Seattle, WA
    98195-1800, USA
2.  School of Zoology, Tel Aviv University, Tel Aviv, Israel
3.  Department of Ecology, Evolution, and Organismal Biology, Iowa State University, 
Ames, IA 50011-4009, USA
Corresponding Author: Lauren Buckley,
[lbuckley\@uw.edu](mailto:lbuckley@uw.edu){.email}

@@

1.  Much understanding of responses to climate change and variability
    relies on the assumption that organismal body temperatures are equal
    to temporally averaged air temperatures high above the ground.
    However, most organisms live near the ground and acute exposure to
    solar and thermal radiation and thermal extremes can substantially elevate their
    body temperatures.
2.  We introduce the TrenchR package, which aids in Translating
    Environmental Change into organismal responses. The package includes
    microclimate models to vertically scale weather station data to
    organismal heights. The biophysical modelling tools include both
    general models for heat flows and specific models to predict body
    temperatures for a variety of ectothermic taxa. Additional functions
    model and temporally partition air and soil temperatures and solar
    radiation. We offer utility functions to aid in estimating the
    organismal and environonmental parameters needed for biophysical
    ecology.
3.  TrenchR focuses on simple and modular functions so users
    can create transparent and flexible models for biophysical applications.
4.  The package aims to introduce and enable microclimate and
    biophysical modelling to improve ecological and evolutionary
    forecasting. We further this aim through a series of educational
    modules to introduce the field of biophysical ecology.

@@

Responses of organisms and ecosystems to climate change are heterogeneous and thus
inconsistent with current predictive models [@maguire2015modeling]. Some
of these predictive shortcomings stem from omitting spatial and temporal
environmental variation and how it interacts with organismal phenotypes
(e.g., size, coloration) [@nadeau2017coarse; @potter2013microclimatic].
Many analyses assume that organismal body temperatures are equal to air
temperatures at weather station height (usually 2 meters). However, air
temperatures near the ground where most organisms reside can be
considerably warmer, and absorption of solar and thermal radiation can raise body
temperatures well above air temperatures, resulting in potential
discrepancies of tens of degrees Celsius [@sinclair2016; @sunday2014;
@kearney2009]. Applications are increasingly demonstrating the
importance of accounting for microclimate variation
[@bramer2018advances; @butikofer2020problem; @lembrechts2019comparing].
Acute thermal stress events are often more relevant to the physiology
and energy balance of organisms than daily, monthly, or annually
averaged environmental conditions. Yet, daily and seasonal
environmental variation is often neglected [@dillon2016introduction;
@dillon2016life]. We aim to help remedy these shortcomings by
introducing the TrenchR package
(<https://github.com/trenchproject/TrenchR>) as a component of the
TrEnCh project (<https:www.trenchproject.com>), which builds
computational and visualization tools to Translate Environmental Change
into organismal responses.

@@

The TrenchR package is intended to complement
the NicheMapR package, which includes sophisticated microclimate
[@kearney2017nichemapr] and biophysical [@kearney2020nichemapr] models.
Although the Fortran source code was recently released, the complexity
of NicheMapR functions can make it difficult to understand and modify the
source code. Many of the approaches are similar between the packages,
but TrenchR provides simple functions to aid the understanding and
transparency of biophysical approaches. The TrenchR functions are
modular and easy adapt to a variety of organisms and research
questions. Although many of the TrenchR functions are general, some
components of TrenchR are best suited for ectothermic animals. We have limited inclusion of plant specific biophysical models since
they are the focus of several R packages [tealeaves: [@muir2019tea];
plantecophys: [@duursma2015plantecophys]]. Some energy budget
calculation components of biophysical models are also available in the
ThermImage package [@tattersall2017]. Microclimate models, with an
emphasis on describing spatial variation, are also available in the
Microclima package [@maclean2019microclima].

@@

The TrenchR package (<http://trenchproject.github.io/TrenchR/>) aims to promote transparency and reproducibility.
We welcome contributions and corrections from users. Our package is
built using the devtools methodology
(<https://github.com/hadley/devtools>) in Github. Issues and feature
requests can be contributed in Github
(<https://github.com/trenchproject/TrenchR/>).

Functions were adapted from biophysical ecology textbooks
[@campbell2000introduction; @gates1980biophysical] and research
articles. The source reference is included in each function's header. We
have incorporated default parameters (e.g., organism emissivity or heat
conduction rates) when general values are available. Many functions
additionally include comments that describe potential parameterizations.
The biophysical ecology texts [@campbell2000introduction;
@gates1980biophysical] include many tables with parameter values and we
provide several data tables in TrenchR.

@@

### Vignettes and introductory modules

@@

For additional background on microclimate and biophysical modelling, we
have updated a series of educational modules entitled 
*Physical Processes in Ecosystems*. We intend the modules to provide a more
contained and accessible introduction to microclimate and biophysical
modelling than that included in classic textbooks
[@campbell2000introduction; @gates1980biophysical]. The modules provide
less detailed coverage of many of the topics contained in the books. The
series of 14 modules starts with an overview of the calculus and physics
principles underlying the modelling. Thermodynamics and energy budget
modelling are then introduced with detailed examples. Modules address
the climate space concept, operative temperatures, biophysical models
for leaves and sheep, and heat flow in soils. We have revised the
modules (which originated from an NSF training grant in 1979 and lacked
broad distribution) to include R code and utilization of TrenchR
functions. We expanded the original series of tutorials to include
introductions by Michael Kearney to the Microclim environmental data
[@kearney2014microclim] and integrating the Microclim data with
biophysical modelling to examine broad scale climatic limits. The
modules are provided in html form and available for download as a pdf at
the bookdown server
(<https://bookdown.org/huckley/Physical_Processes_In_Ecosystems/>). The
R markdown files are available in Github (<https://github.com/trenchproject/TrenchRmodules>).

@@

We illustrate the use of the TrenchR package by estimating an energy
budget for a *Sceloporus* lizard on June 1, 2020 in Santa Fe, New
Mexico, USA (35.69, -105.944, elevation: 2121 meters). The example is also
incorporated in the *Using energy balances to estimate body temperatures* 
vignette. We start by generating environmental inputs. Using these inputs,
we estimate the energy budget with component functions. Finally, we use 
an integrated biophysical model to estimate operative environmental temperatures.  
See Appendix S1 for another example examining a
time series of microclimate data and applying and testing the
biophysical models.

@@

We first prepare the environmental data for analysis. We will estimate 
hourly air and soil temperatures and radiation usinf a function describing diurnal temperature variation. We start by
estimating the day of year and the timing of sunrise and sunset:

```{r, echo=FALSE}
library(TrenchR)

#to be REMOVED
Tb_CampbellNorman=function(T_a, T_g, S, alpha_L=0.96, epsilon=0.96, c_p=29.3, D, V){

  stopifnot(T_a>200, T_a<400, epsilon>=0.5, epsilon<=1, c_p>=0, D>0, V>=0)

  #Stefan-Boltzmann constant
  sigma= 5.673*10^(-8) #W m^(-2) K^(-4)

  #solar and thermal radiation absorbed
  L_a=sigma*T_a^4  # (10.7) long wave flux densities from atmosphere 
  L_g=sigma*T_g^4  # (10.7) long wave flux densities from ground
  F_a=0.5; F_g=0.5 #proportion of organism exposure to air and ground, respectively
  R_abs= S+alpha_L*(F_a*L_a+F_g*L_g) # (11.14) Absorbed radiation

  #thermal radiation emitted
  Qemit= epsilon*sigma*T_a^4

  #conductance
  g_Ha=1.4*0.135*sqrt(V/D) # boundary conductance, factor of 1.4 to account for increased convection (Mitchell 1976), assumes forced conduction
  g_r= 4*epsilon*sigma*T_a^3/c_p # (12.7) radiative conductance

  # operative environmental temperature
  T_e=T_a+(R_abs-Qemit)/(c_p*(g_r+g_Ha))

  return(T_e) 
}
```

@@

doy = day_of_year("2020-06-01", format= "%Y-%m-%d") # Find Julian Date (DOY)
snoon = solar_noon(lon = lon, doy = doy) # Estimate solar noon
dayl = daylength(lat = lat, doy = doy) # Estimate day length
tr = snoon - dayl / 2 # Time of sunrise
ts = snoon + dayl / 2 # Time of sunset
```

@@

```{r}
# Estimate zenith angle (degrees)
psi_deg = sapply(1:24, FUN = zenith_angle, doy = doy, lat = lat, lon = lon) 
# Convert to radians
psi_rad = degree_to_radian(psi_deg) 
# Estimate radiation
Srad = sapply(psi_rad, FUN = estimate_radiation, doy = doy, tau = tau, elev = elev, rho = rho)

@@

# Estimate surface area and sihouette area
A = surface_area_from_mass(mass, "lizard") # surface area (m^2)
psa = sapply(psi_deg, proportion_silhouette_area, taxon = "lizard", posture = "prostrate") # Projected silhouette area (m^2)

# Change negative values to zero
psa[psa < 0] = 0
```

@@

# Estimate the heat transfer coefficient using an empirical relationship for lizards
H_L = heat_transfer_coefficient(V = V_liz, D = svl, K = 25.7 * 10^(-3), nu = 15.3 * 10^(-6), taxa = "lizard_surface")

# Estimate the heat transfer coefficient using a spherical approximation
H_L2= heat_transfer_coefficient_approximation(V = V_liz, D = svl, K = 25.7 * 10^(-3), nu = 15.3 * 10^(-6), taxa = "lizard")

@@

The microclimate models indicate that air temperatures at lizard height
are similar to surface temperatures (Figure 1). The biophysical models
indicate that solar radiation will elevate lizard body temperatures far
above air temperatures and that the lizard will face thermal stress if
it is unable to seek shade (Figure 1). The three biophysical models
predict different body temperatures during peak period of solar
radiation because they model interactions with the ground differently
and users are encouraged to review the details of each biophysical model
before selection. Differences in estimated body temperatures are
accentuated by the high level of solar radiation.

@@

**Figure 1**. Body temperatures (Te) are predicted to drastically exceed
air temperature when lizards are exposed to high levels of solar
radiation. Air temperatures (Ta, °C) at lizard height (0.02m) are
predicted to exceed air temperatures at 2m and to be similar to surface
temperatures (Ts). We estimate body temperatures using two general
energy budgets [solid: Tb_Gates(); dotted: Tb_CampbellNorman()] and a
lizard specific biophysical model [dashed: Tb_lizard()] that differ in
how they model heat exchanges.

@@

## Acknowledgements
We thank Ray Huey, Joel Kingsolver, Luke Miller, Warren
Porter, and Art Woods for assistance and input and Brian Helmuth for providing functions. We particularly thank
Michael Kearney for extensive input on TrenchR and its alignment with
NicheMapR and for contributing two educational modules.

@@

LB conceived and designed the package and lead all aspect of the
packages creation and dissemination; BBO, IC, AJ, OL, AM, ER, YS wrote and
edited functions and associated package resources; AJ configured the
package; AM, BBO, YS, and LB converted and edited the educational
modules. All authors contributed critically to the drafts and gave final
approval for publication.

>>>>

---
title: 'TrenchR: an R package for transparent environmental and ecological biophysics'
author: "Lauren B. Buckley, Bryan A. Briones Ortiz, Isaac Caruso, Aji John, Ofir Levy, Abigail V. Meyer, Eric A. Riddell, Yutaro Sakairi, and Juniper Simonis"
output:
  html_document: default
  word_document: default
  pdf_document: default
bibliography: bibliography.bib
csl: methods-in-ecology-and-evolution.csl
editor_options: 
  markdown: 
    wrap: 72
---

@@

Lauren B. Buckley^1^, Bryan A. Briones Ortiz^1^, Isaac Caruso^1^, Aji
John^1^, Ofir Levy^2^, Abigail V. Meyer^1^, Eric A. Riddell^3^, Yutaro
Sakairi^1^, and Juniper L. Simonis^4^

1.  Department of Biology, University of Washington, Seattle, WA
    98195-1800, USA
2.  School of Zoology, Tel Aviv University, Tel Aviv, Israel
3.  Department of Ecology, Evolution, and Organismal Biology, Iowa State
    University, Ames, IA 50011-4009, USA
4.  DAPPER Stats, 3519 NE 15th Avenue, Suite 467, Portland, Oregon
    97212, USA

Corresponding Author: Lauren Buckley,
[lbuckley\@uw.edu](mailto:lbuckley@uw.edu){.email}

@@

1.  Much understanding of organimsal responses to climate change and
    variability relies on the assumption that body temperatures are
    equal to temporally averaged air temperatures high above the ground.
    However, most organisms live near the ground and acute exposure to
    solar and thermal radiation and thermal extremes can substantially
    elevate their body temperatures.
2.  We introduce the TrenchR package, which aids in Translating
    Environmental Change into organismal responses. The package includes
    microclimate models to vertically scale weather station data to
    organismal heights. The biophysical modelling tools include both
    general models for heat flows and specific models to predict body
    temperatures for a variety of ectothermic taxa. Additional functions
    model and temporally partition air and soil temperatures and solar
    radiation. We offer utility functions to aid in estimating the
    organismal and environonmental parameters needed for biophysical
    ecology.
3.  TrenchR focuses on simple and modular functions so users can create
    transparent and flexible models for biophysical applications.
4.  The package aims to introduce and enable microclimate and
    biophysical modelling to improve ecological and evolutionary
    forecasting. We further this aim through a series of educational
    modules to introduce the field of biophysical ecology.

@@

Responses of organisms and ecosystems to climate change are
heterogeneous and thus inconsistent with current predictive models
[@maguire2015modeling]. Some of these predictive shortcomings stem from
omitting spatial and temporal environmental variation and how it
interacts with organismal phenotypes (e.g., size, coloration)
[@nadeau2017coarse; @potter2013microclimatic]. Many analyses assume that
organismal body temperatures are equal to air temperatures at weather
station height (usually 2 meters). However, air temperatures near the
ground where most organisms reside are often considerably warmer, and
absorption of solar and thermal radiation can raise body temperatures
well above air temperatures, resulting in potential discrepancies of
tens of degrees Celsius [@sinclair2016; @sunday2014; @kearney2009].
Applications are increasingly demonstrating the importance of accounting
for microclimate variation [@bramer2018advances; @butikofer2020problem;
@lembrechts2019comparing]. Acute thermal stress events are often more
relevant to the physiology and energy balance of organisms than daily,
monthly, or annually averaged environmental conditions. Yet, daily and
seasonal environmental variation is often neglected
[@dillon2016introduction; @dillon2016life]. We aim to help remedy these
shortcomings by introducing the TrenchR package
(<https://github.com/trenchproject/TrenchR>) as a component of the
TrEnCh project (<https:www.trenchproject.com>), which builds
computational and visualization tools to Translate Environmental Change
into organismal responses.

@@

Estimating how organisms experience their thermal environment generally
entails two classes of models [@campbell2000introduction;
@gates1980biophysical] (Figure 1). Microclimate models allow scaling conditions
from sensors to organism height. They can characterize heat and air
transport to estimate vertical air and soil temperature and wind
profiles. They can also characterize direct and diffuse solar radiation
and longwave radiation emanating from the sky, ground, and surrounding
objects like vegetation. Biophysical models estimate body temperatures
using an energy budget to balance heat exchanges between organisms and
their environment.

```{r, echo=FALSE, fig.height=4, out.width = "75%"}
knitr::include_graphics('figures/TrenchRBiophys.png')
```

**Figure 1** A) Microclimate models scale air and soil temperatures and windspeeds from sensor to organismal heights. B) Biophysical models balance heat exchanges between organisms and their environment to estimate body temperatures [@porter1973behavioral]. Grasshopper body temperatures vary substantially from vegetation, air, and ground temperatures due to heat exchange.

The TrenchR package is intended to complement the NicheMapR package,
which includes sophisticated microclimate [@kearney2017nichemapr] and
biophysical [@kearney2020nichemapr] models. Although the Fortran source
code was recently released, the complexity of NicheMapR functions can
make it difficult to understand and modify the source code. Many of the
approaches are similar between the packages, but TrenchR provides simple
functions to aid the understanding and transparency of biophysical
approaches. The TrenchR functions are modular and easily adapt to a
variety of organisms and research questions. Although many of the
TrenchR functions are general, some components of TrenchR are best
suited for ectothermic animals. We omit plant specific biophysical
models since they are the focus of several R packages [tealeaves:
[@muir2019tea]; plantecophys: [@duursma2015plantecophys]]. Some energy
budget calculation components of biophysical models are also available
in the ThermImage package [@tattersall2017]. Microclimate models, with
an emphasis on describing spatial variation, are also available in the
Microclima package [@maclean2019microclima].

@@

The TrenchR package (<http://trenchproject.github.io/TrenchR/>) aims to
promote transparency and reproducibility. We welcome contributions and
corrections from users. Our package is built using the devtools
methodology (<https://github.com/r-lib/devtools>) in Github. Issues and
feature requests can be contributed in Github
(<https://github.com/trenchproject/TrenchR>). The packageis available
via Github and has been submitted to CRAN.

We adapted functions from biophysical ecology textbooks
[@campbell2000introduction; @gates1980biophysical] and research
articles. Source references are included in function headers. We have
incorporated default parameters (e.g., organism emissivity or heat
conduction rates) when general values are available. Many functions
additionally include comments that describe potential parameterizations.
The biophysical ecology texts [@campbell2000introduction;
@gates1980biophysical] include many tables with parameter values and we
provide several data tables in TrenchR.

@@

### Vignettes and introductory tutorials

@@

For additional background on microclimate and biophysical modelling, we
have updated a series of tutorials entitled *Physical Processes in
Ecosystems*. We intend the tutorials to provide a more contained and
accessible introduction to microclimate and biophysical modelling than
that included in classic textbooks [@campbell2000introduction;
@gates1980biophysical]. The tutorials provide less detailed coverage of
many of the topics contained in the books. The series of 14 tutorials
starts with an overview of the calculus and physics principles
underlying the modelling. Thermodynamics and energy budget modelling are
then introduced with detailed examples. Tutorials address the climate
space concept, operative temperatures, biophysical models for leaves and
sheep, and heat flow in soils. We have revised the tutorials (which
originated from an NSF training grant in 1979 and lacked broad
distribution) to include R code and utilization of TrenchR functions. We
expanded the original series to include tutorials contributed by M.
Kearney introducing the Microclim environmental data
[@kearney2014microclim] and integrating the Microclim data with
biophysical modelling to examine broad scale climatic limits. The
tutorials are provided in html form and available for download as a pdf
at the bookdown server
(<https://bookdown.org/huckley/Physical_Processes_In_Ecosystems/>). The
R markdown files are available in Github
(<https://github.com/trenchproject/TrenchRmodules>).

@@

We illustrate the use of the TrenchR package by estimating an energy
budget for a *Sceloporus* lizard on June 1, 2021 in Santa Fe, New
Mexico, USA (35.69, -105.944, elevation: 2121 meters). The example is
also incorporated in the *Using energy balances to estimate body
temperatures* vignette. We start by generating environmental inputs.
Using these inputs, we estimate the energy budget with component
functions. Finally, we use an integrated biophysical model to estimate
operative environmental temperatures. See Appendix S1 for another
example examining a time series of microclimate data and applying and
testing the biophysical models.

@@

We first prepare the environmental data for analysis. We will estimate
hourly air and soil temperatures and radiation using a function
describing diurnal temperature variation. We start by estimating the day
of year and the timing of sunrise and sunset:

```{r, echo=FALSE}
library(TrenchR)
```

@@

doy = day_of_year("2021-06-01", format= "%Y-%m-%d") # Find Julian Date (DOY)
snoon = solar_noon(lon = lon, doy = doy) # Estimate solar noon
dayl = daylength(lat = lat, doy = doy) # Estimate day length
tr = snoon - dayl / 2 # Time of sunrise
ts = snoon + dayl / 2 # Time of sunset
```

@@

```{r}
# Estimate zenith angle (degrees)
psi_deg = sapply(1:24, FUN = zenith_angle, doy = doy, lat = lat, lon = lon) 
# Convert to radians
psi_rad = degrees_to_radians(psi_deg) 
# Estimate radiation
Srad = sapply(psi_rad, FUN = solar_radiation, doy = doy, tau = tau, elev = elev, rho = rho)

@@

# Estimate surface area and the proportion sihouette area
A = surface_area_from_mass(mass, "lizard") # surface area (m^2)
psa = sapply(psi_deg, proportion_silhouette_area, taxon = "lizard", posture = "prostrate") 
# Change negative values to zero
psa[psa < 0] = 0
```

@@

# Estimate the heat transfer coefficient using an empirical relationship for lizards
H_L = heat_transfer_coefficient(V = V_liz, D = svl, K = 25.7 * 10^(-3), nu = 15.3 * 10^(-6), taxon = "lizard_surface")

# Estimate the heat transfer coefficient using a spherical approximation
H_L2= heat_transfer_coefficient_approximation(V = V_liz, D = svl, K = 25.7 * 10^(-3), nu = 15.3 * 10^(-6), taxon = "lizard")

@@

The microclimate models indicate that air temperatures at lizard height
are similar to surface temperatures (Figure 2). The biophysical models
indicate that solar radiation will elevate lizard body temperatures far
above air temperatures and that the lizard will face thermal stress if
it is unable to seek shade (Figure 2). The three biophysical models
predict different body temperatures during peak period of solar
radiation because they model interactions with the ground differently
and users are encouraged to review the details of each biophysical model
before selection. Differences in estimated body temperatures are
accentuated by the high level of solar radiation.

@@

**Figure 2**. Body temperatures (Te) are predicted to drastically exceed
air temperature when lizards are exposed to high levels of solar
radiation. Air temperatures (Ta, °C) at lizard height (0.02m) are
predicted to exceed air temperatures at 2m and to be similar to surface
temperatures (Ts). We estimate body temperatures using two general
energy budgets [solid: Tb_Gates(); dotted: Tb_CampbellNorman()] and a
lizard specific biophysical model [dashed: Tb_lizard()] that differ in
how they model heat exchanges.

@@

## Acknowledgements

We thank A. Cannistra, R. Huey, J. Kingsolver, L. Miller, W. Porter, and
H. A. Woods for assistance and input and B. Helmuth for providing
functions. We particularly thank M. Kearney for extensive input on
TrenchR and its alignment with NicheMapR and for contributing two
tutorials.

@@

LB conceived and designed the package and lead all aspect of the
packages creation and dissemination; BBO, IC, AJ, OL, AM, ER, YS, and JS
wrote and edited functions and associated package resources; AJ and JS
configured the package; AM, BBO, YS, and LB converted and edited the
educational modules. All authors contributed critically to the drafts
and gave final approval for publication.
